---
title: "Project proposal"
author: "Team Hull"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(here)
```
```{r}
Substance_Abuse <- read.csv(here::here("data/Substance Abuse.csv"))
```
```{r}
Weight_Issues <- read.csv(here::here("data/Weight Issues.csv"))
```




test

## 1. Introduction

The data we are going to be using is "Youth Risk Behavior Surveillance System (YRBSS): Middle School". This data has a wide range of variables and entries from 1991-2017 that monitored young peoples behaviors (violence; tobacco use; substance abuse; sexual health; dietary health; and physical health). These data sets specifically focuses on substance abuse levels and dietary Health over a 26 year period within middle school.

This data was collected by the Centers for Disease Control and Prevention(CDC). As of 25/10/2021, this data was last updated on March 29th, 2019. 

The 35 variables in this data set are year, loactionabbr, LocationDesc, DataSource, Topic, Subtopic, ShortQuestionText, Greater_Risk_Question, Description, Data_Value_Symbol, Data_Value_Type, Greater_Risk_Data_Value, Greater_Risk_Data_Value_Footnote_Symbol, Greater_Risk_Data_Value_Footnote, Greater_Risk_Low_Confidence_Limit, Lesser_Risk_Question,
Lesser_Risk_Data_Value, Lesser_Risk_Data_Value_Footnote_Symbol, Lesser_Risk_Data_Value_Footnote, Lesser_Risk_Low_Confidence_Limit, Lesser_Risk_High_Confidence_Limit, Sample_Size, Sex, Race, Grade, GeoLocation, TopicId, SubTopicID, QuestionCode, LocationId, StratID1, StratID2, StratID3 and StratificationType.

The question we will be looking at is the change in substance abuse and weight issues across the span of 26 years, from 1991-2017, in middle school students and what variables, such as race or sex, or grade, is more susceptible to theses issues.

## 2. Data

```{r Data-frame, message = FALSE}
glimpse(Substance_Abuse)
glimpse(Weight_Issues)
```

## 3. Data analysis plan

The outcome (response, Y) and predictor (explanatory, X) variables you will use to answer your question.
The comparison groups you will use, if applicable.
Very preliminary exploratory data analysis, including some summary statistics. and visualizations, along with some explanation on how they help you learn more about your data. (You can add to these later as you work on your project.)
The statistical method(s) that you believe will be useful in answering your question(s). (You can update these later as you work on your project.)
What results from these specific statistical methods are needed to support your hypothesized answer?

The Variables that will be used to answer the question will be mainly year and and subtopic to see whether there is some sort of trend with the way these issues develop as high schoolers complete each grade level. Furthermore, to see the breakdown on who is affected more, sex, race, and grade will be used in comparison. This will help us see if there are any trends, such as more people starting to abuse substances as the approach grade x, or if male or female high school children are more prone to having behavioral issues with weight and dietary health. 

After taking a preliminary look at the dataset, we can see that the instances of substance abuse and weight probably start to increase more sharply from 10th grade onward. 9th grade seems to be relatively tame with regards to these things, and each subsequent year seems to grow in problems. We could probably look to statistical methods such as calculating the mean year at which students in particular regions are experiencing these problems. We could also map frequency of instances of substance abuse or weight issues by population, in order to see whether, for example, certain urban/rural areas or more conservative/liberal states are experiencing more or less of each of these problems. 

##4 

```{r}
Substance_Abuse %>%
  group_by(Sex, Grade) %>%
  summarise(n())
```

```{r}
Substance_Abuse %>%
  group_by(Sex, Grade) %>%
  summarise(n())


```


```{r}
Substance_Abuse %>%
  group_by(Sex) %>%
  ggplot(aes(x=Grade, fill=Sex)) +
  geom_bar(stat="count")
```
This graph shows the amount of each grade that is represented on this list of Substance Abuse related events grouping by Sex. This shows a fairly even split between male and females.
```{r}
Weight_Issues %>%
  group_by(Sex) %>%
  ggplot(aes(x=Greater_Risk_Question, fill=Sex)) +
  geom_bar()
```
This graph shows the ammount of each grade represnted in the data for weight issues and shows there is no great disparity between races on the issue.

