---
title: "Project proposal"
author: "Team name"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

*For instructions on what each section should include, please see the [project page](https://idsed.digital/assessments/project/#proposal) on the course website. Remove this text when completing your proposal*.

test

## 1. Introduction

The data we are going to be using is "Youth Risk Behavior Surveillance System (YRBSS): Middle School". This data has a wide range of variables and entries from 1991-2017 that monitored young peoples behaviors (violence; tobacco use; substance abuse; sexual health; dietary health; and physical health). These data sets specifically focuses on substance abuse levels and dietary Health over a 26 year period within middle school.

This data was collected by the Centers for Disease Control and Prevention(CDC). As of 25/10/2021, this data was last updated on March 29th, 2019. 

The 35 variables in this data set are year, loactionabbr, LocationDesc, DataSource, Topic, Subtopic, ShortQuestionText, Greater_Risk_Question, Description, Data_Value_Symbol, Data_Value_Type, Greater_Risk_Data_Value, Greater_Risk_Data_Value_Footnote_Symbol, Greater_Risk_Data_Value_Footnote, Greater_Risk_Low_Confidence_Limit, Lesser_Risk_Question,
Lesser_Risk_Data_Value, Lesser_Risk_Data_Value_Footnote_Symbol, Lesser_Risk_Data_Value_Footnote, Lesser_Risk_Low_Confidence_Limit, Lesser_Risk_High_Confidence_Limit, Sample_Size, Sex, Race, Grade, GeoLocation, TopicId, SubTopicID, QuestionCode, LocationId, StratID1, StratID2, StratID3 and StratificationType.

The question we will be looking at is the change in substance abuse and weight issues across the span of 26 years, from 1991-2017, in middle school students and what variables, such as race or sex, or grade, is more susceptible to theses issues.

## 2. Data

```{r Data-frame, message = FALSE}
glimpse(Substance_Abuse)
glimpse(Weight_Issues)
```

## 3. Data analysis plan



