---
title: "Youth Risk Behavior"
subtitle: <font size='6'>Substance Abuse & Weight issues</font>
author: <font size='8'>Team Hull <br> Philip Wieczorek, Andrew Allan, Joshua Brown, Fengshou Miao</font>
institute: <font size='10'>University of Edinburgh</font>
date: <font size = '6'>`r Sys.Date()`</font>
output:
  revealjs::revealjs_presentation:
    self_contained: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(rmarkdown)
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(revealjs)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
Substance_Abuse <- read.csv(here::here("data/Substance Abuse.csv"))
Weight_Issues <- read.csv(here::here("data/Weight Issues.csv"))
```

```{r echo=FALSE}
Substance_Abuse_Edited <- Substance_Abuse %>%
  mutate(Pop_at_Grisk = (Greater_Risk_Data_Value/100) * Sample_Size)%>%
  group_by(YEAR, LocationDesc, Sample_Size, Pop_at_Grisk, Sex, Race, Grade, Greater_Risk_Data_Value,Greater_Risk_High_Confidence_Limit, Greater_Risk_Low_Confidence_Limit, Lesser_Risk_Low_Confidence_Limit, Lesser_Risk_High_Confidence_Limit, GeoLocation, QuestionCode, LocationId)
```

```{r echo=FALSE}
Weight_Issues_Edited <- Weight_Issues %>%
  mutate(Pop_at_Grisk = (Greater_Risk_Data_Value/100) * Sample_Size)%>%
  group_by(YEAR, LocationDesc, Sample_Size, Pop_at_Grisk, Sex, Race, Grade, Greater_Risk_Data_Value, Greater_Risk_High_Confidence_Limit, Greater_Risk_Low_Confidence_Limit, Lesser_Risk_High_Confidence_Limit, Lesser_Risk_Low_Confidence_Limit, GeoLocation, QuestionCode, LocationId)
```

## Our goals with this project were to...

- Discover whether risky behavior such as "substance abuse" and "weight issues" may appear more or less frequently among different categorical variables such as sex, race, etc. 

- Model how risky behavior trends throughout different grades

- Map important or interesting trends that we found along the way


# Our Dataset: Substance Abuse

```{r Substance_Abuse_Edited Glimpse, echo=FALSE}
glimpse(Substance_Abuse_Edited)

```

# Our Dataset: Weight Issues

```{r Weight_Issues_Glimpse, echo=FALSE}
glimpse(Weight_Issues_Edited)

```
# Substance Abuse over Time

```{r SA_Time, echo=FALSE}
SAE2 <- Substance_Abuse_Edited %>%
  filter(
    Sex != "Total",
    Greater_Risk_Question == "Drank alcohol for the first time before age 11 years"
  ) %>%
  drop_na(Greater_Risk_Data_Value) %>%
  group_by(Sex, YEAR) %>%
  summarise(Percentage_Greater_Risk = weighted.mean(Greater_Risk_Data_Value, Sample_Size)) 

SAE2 %>%
  ggplot(aes(x = YEAR, y = Percentage_Greater_Risk, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal()

```


# Substance Abuse over Time

- Females at lower risk overall than men to abuse substances

- Risk of substance abuse has been going down over time
  - Although there was a slight resurgence in the latest year collected
  
# Weight Issues and Geography

```{r WI_Geo, echo = FALSE}
Weight_Issues_Edited %>%
  group_by(LocationDesc) %>%
  drop_na(Greater_Risk_Data_Value) %>%
  summarise(bingo = weighted.mean(Greater_Risk_Data_Value, Sample_Size)) %>%
  ggplot(aes(x = reorder(LocationDesc,-bingo) , y = bingo, fill = LocationDesc)) +
  geom_bar(stat = "identity", position = "dodge") +
theme(axis.text.x = element_text(angle = 90, size = 5), legend.position = "None") +
xlab("Location") +
ylab("% of the Population at Greater Risk")
```

# Weight Issues and Geography

- Overall, USA has a similar incidence of "weight issues" across the board

- At first glance, seems as if coasts (generally but not always more urban) exhibit more weight issues

- Some spread of rural across the board, a number of islands s.a. Northern Mariana Islands and especially American Samoa trending towards fewer weight issues
  - This could have to do with culture; heavier weight more accepted/admired in Samoan culture
  
# Percentage Describing Themselves as Overweight

```{r Overweight_Percent, echo=FALSE}
Weight_Issues %>%
  filter(
    Greater_Risk_Question == "Described themselves as slightly or very overweight",
    Sex != "Total"
  ) %>%
  drop_na(Greater_Risk_Data_Value) %>%
  group_by(Sex, YEAR) %>%
  summarise(Described_as_overweight = weighted.mean(Greater_Risk_Data_Value, Sample_Size)) %>%
  ggplot(aes(x = as.character(YEAR), y = Described_as_overweight, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  xlab("Year") +
  ylab("% Describing Themselves as Overweight")
```

## reveal.js

- The presentation is created using the `reveal.js` package

- You can create a slide show broken up into sections by using the # and ## heading tags

- You can also create a new slide without a header using a horizontal rule (----)

----

- You can also add pauses between content on a slide using (". . .")

. . .

Like this.

...or using "::: fragment"

::: fragment

Like this.

:::

# Layouts

You can use plain text

- or bullet points

<div>
  <div style="float: left; width: 50%;">or text in two columns</div>
  <div style="float: left; width: 50%;">like this</div>
</div>

because you can combine HTML and Markdown syntax.

<p style="font-size: 8pt; position: absolute; bottom: -5em; width: 100%; text-align: center;">*You can also add footnotes</p>

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

## Plot and text

::: columns
:::: column
* Some text
* goes here
::::

:::: column
```{r warning=FALSE,  echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
```
::::
:::

## Different widths

::: columns
:::: {.column width="40%"}
* Some text
* goes here
::::

:::: {.column width="60%"}
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
```
::::
:::

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
```

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

## {data-background-image="https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80"}

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs (\$...\$), e.g. $\alpha+\beta$. You can use the display style with double dollar signs:

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

# Feeling adventurous?

You can learn more about the options available at the following pages:

- https://github.com/rstudio/revealjs

- https://bookdown.org/yihui/rmarkdown/revealjs.html

- http://www.jenniferbradham.org/post/reveal_js_markdown/

- https://revealjs.com/