---
title: "Youth Risk Behavior"
subtitle: <font size='6'>Substance Abuse & Weight issues</font>
author: <font size='8'>Team Hull <br> Philip Wieczorek, Andrew Allan, Joshua Brown, Fengshou Miao</font>
institute: <font size='10'>University of Edinburgh</font>
date: <font size = '6'>`r Sys.Date()`</font>
output:
  revealjs::revealjs_presentation:
    self_contained: false
    theme: serif
    highlight: pygments
    center: true
---

---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(rmarkdown)
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(revealjs)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
Substance_Abuse <- read.csv(here::here("data/Substance Abuse.csv"))
Weight_Issues <- read.csv(here::here("data/Weight Issues.csv"))
```

```{r echo=FALSE}
Substance_Abuse_Edited <- Substance_Abuse %>%
  mutate(Pop_at_Grisk = (Greater_Risk_Data_Value/100) * Sample_Size)%>%
  group_by(YEAR, LocationDesc, Sample_Size, Pop_at_Grisk, Sex, Race, Grade, Greater_Risk_Data_Value,Greater_Risk_High_Confidence_Limit, Greater_Risk_Low_Confidence_Limit, Lesser_Risk_Low_Confidence_Limit, Lesser_Risk_High_Confidence_Limit, GeoLocation, QuestionCode, LocationId)
```

```{r echo=FALSE}
Weight_Issues_Edited <- Weight_Issues %>%
  mutate(Pop_at_Grisk = (Greater_Risk_Data_Value/100) * Sample_Size)%>%
  group_by(YEAR, LocationDesc, Sample_Size, Pop_at_Grisk, Sex, Race, Grade, Greater_Risk_Data_Value, Greater_Risk_High_Confidence_Limit, Greater_Risk_Low_Confidence_Limit, Lesser_Risk_High_Confidence_Limit, Lesser_Risk_Low_Confidence_Limit, GeoLocation, QuestionCode, LocationId)
```

## Original Ideas

- CDC data from Youth Risk Behavior Surveillance System. created on August 29th 2016 and were last updated March 29th 2019

- Violence; Unprotected sexual practice; Substance abuse; Tobacco use; Weight Issues; Lack of physical activity.

- 2 of the 6 risk behaviors 

## Our goals with this project were to...

- Discover whether risky behavior such as "substance abuse" and "weight issues" may appear more or less frequently among different categorical variables such as sex, race, etc. 

- Model how risky behavior trends throughout different grades

- Map important or interesting trends that we found along the way


# Our Dataset: Substance Abuse

```{r Substance_Abuse_Edited Glimpse, echo=FALSE}
glimpse(Substance_Abuse_Edited)

```

# Our Dataset: Weight Issues

```{r Weight_Issues_Glimpse, echo=FALSE}
glimpse(Weight_Issues_Edited)

```
# Substance Abuse over Time

```{r SA_Time, echo=FALSE}
SAE2 <- Substance_Abuse_Edited %>%
  filter(
    Sex != "Total",
    Greater_Risk_Question == "Drank alcohol for the first time before age 11 years"
  ) %>%
  drop_na(Greater_Risk_Data_Value) %>%
  group_by(Sex, YEAR) %>%
  summarise(Percentage_Greater_Risk = weighted.mean(Greater_Risk_Data_Value, Sample_Size)) 

SAE2 %>%
  ggplot(aes(x = as.character(YEAR), y = Percentage_Greater_Risk, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  ylab("% at Greater Risk of Substance Abuse") +
  xlab("Years")

```


# Substance Abuse over Time

- Females at lower risk overall than men to abuse substances

- Risk of substance abuse has been going down over time
  - Although there was a slight resurgence in the latest year collected
  
# Weight Issues and Geography

```{r WI_Geo, echo = FALSE}
Weight_Issues_Edited %>%
  group_by(LocationDesc) %>%
  drop_na(Greater_Risk_Data_Value) %>%
  summarise(bingo = weighted.mean(Greater_Risk_Data_Value, Sample_Size)) %>%
  ggplot(aes(x = reorder(LocationDesc, -bingo) , y = bingo, fill = LocationDesc)) +
  geom_bar(stat = "identity", position = "dodge") +
theme(axis.text.x = element_text(angle = 90, size = 5), legend.position = "None") +
xlab("Location") +
ylab("% of the Population at Greater Risk")
```

# Weight Issues and Geography

- Overall, USA has a similar incidence of "weight issues" across the board

- At first glance, seems as if coasts (generally but not always more urban) exhibit more weight issues

# Weight Issues and Geography

- Some spread of rural across the board, a number of islands s.a. Northern Mariana Islands and especially American Samoa trending towards fewer weight issues
  
- This could have to do with culture; heavier weight more accepted/admired in Samoan culture
  
# Percentage Describing Themselves as Overweight

```{r Overweight_Percent, echo=FALSE}
Weight_Issues %>%
  filter(
    Greater_Risk_Question == "Described themselves as slightly or very overweight",
    Sex != "Total"
  ) %>%
  drop_na(Greater_Risk_Data_Value) %>%
  group_by(Sex, YEAR) %>%
  summarise(Described_as_overweight = weighted.mean(Greater_Risk_Data_Value, Sample_Size)) %>%
  ggplot(aes(x = as.character(YEAR), y = Described_as_overweight, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  xlab("Year") +
  ylab("% Describing Themselves as Overweight")
```

# Percentage Describing Themselves as Overweight

- Females tend to believe they are overweight more frequently than men

- Females have trended down a bit, but has gone up a little bit in recent years, although remaining below 30%

- Males have maintained between 22% and 25%, not much variation over time

# Substance Abuse Issues and Geography

```{r echo=FALSE}
Substance_Abuse_Edited %>%
  group_by(LocationDesc) %>%
  drop_na(Greater_Risk_Data_Value) %>%
  summarise(bingo = weighted.mean(Greater_Risk_Data_Value, Sample_Size)) %>%
  ggplot(aes(x = reorder(LocationDesc, -bingo) , y = bingo, fill = LocationDesc)) +
  geom_bar(stat = "identity", position = "dodge") +
theme(axis.text.x = element_text(angle = 90, size = 5), legend.position = "None") +
xlab("Location") +
ylab("Percentage of the Population at Greater Risk")
```

# Substance Abuse Issues and Geography

- Dallas and Alabama are the areas with the most substance abuse issues by quite a bit.

- American Samoa is the least
 
- There is a greater range of substance abuse issues in these areas than weight issues.

# Substance Abuse Issues by Race

```{r echo = FALSE}
Substance_Abuse_Edited %>%
  group_by(Race) %>%
  drop_na(Greater_Risk_Data_Value) %>%
  summarise(bingo = weighted.mean(Greater_Risk_Data_Value, Sample_Size)) %>%
  ggplot(aes(x = reorder(Race, -bingo) , y = bingo, fill = Race)) +
  geom_bar(stat = "identity", position = "dodge") +
theme(axis.text.x = element_blank()) +
xlab("Race") +
ylab("Percentage of the Population at Greater Risk")
```

# Substance Abuse Issues by Race

- Native American or Other Pacific Islander, Hispanic or Latino and Black or African American have the highest percentage of their populations with substance abuse issues. 

- The race with the lowest percentage of substance abuse issues is Asian.

# What Could We Have Included?

- Comparing common trends between the two datasets (Is geography correlated?)

- Comparing all 6 risk factors in some way. or comparing the 2 risk factors over different school levels.

# What Could We Have Included?

- Looking for more opportunities to find correlation, regression, other trends.

- Perhaps creating a literal map with a "heat map" with an additional package.

- Would give us slightly more precise visualization with respect to geography.

# Data Collection

Our data was collected from the Youth Risk Behavior Surveillance System(YRBSS), for the Centers for Disease Control and Prevention (CDC) and can be vied on the CDCs website.

- Females have greater incidence of "weight issues"

- Males more likely to be exposed to "substance abuse"

# Data Collection

- Some progress made in both, but still plaguing our generation

- Important topic to discuss --> societal relevance, mapping the impact of body positivity and substance abuse reduction campaigns


